---
output: 
  github_document:
    pandoc_args: --webtex
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T)
```

# familial

[![R-CMD-check](https://github.com/ryan-thompson/familial/workflows/R-CMD-check/badge.svg)](https://github.com/ryan-thompson/familial/actions)
[![codecov](https://codecov.io/gh/ryan-thompson/familial/branch/master/graph/badge.svg)](https://github.com/ryan-thompson/familial/actions)

## Overview

An R package for testing familial hypotheses. A paper describing familial hypothesis is forthcoming. Briefly, they are statements of the form:
$$
\begin{aligned}
&\mathrm{H}_0:\mu(\lambda)=\mu_0\text{ for some }\lambda\in\Lambda \\
&\mathrm{H}_1:\mu(\lambda)\neq\mu_0\text{ for all }\lambda\in\Lambda,
\end{aligned}
$$
where $\{\mu(\lambda):\lambda\in\Lambda\}$ is a family of population parameters, e.g., different centers of a distribution induced by the Huber loss function with parameter $\lambda$. In contrast to classic statistical tests such as the $t$ or sign test for the mean or median, familial tests do not require the user to choose a specific version of center, making them quite robust. See the package [vignette](vignette.pdf) for illustrations.

Presently, `familial` supports tests of center via the Huber or the trimmed mean families of location parameters. Both of these families include the mean and median. Testing is carried out using a Bayesian approach whereby the posterior probabilities of the competing hypotheses $\mathrm{H}_0$ and $\mathrm{H}_1$ are computed using the Bayesian bootstrap. One- and two-sample tests are supported, as are directional tests. Methods for visualizing output are provided.

## Installation

<!-- To install the latest stable version from CRAN, run the following code: -->
<!-- ``` {r, eval = F} -->
<!-- install.packages('familial') -->
<!-- ``` -->

To install the latest development version from GitHub, run the following code:
``` {r, eval = F}
devtools::install_github('ryan-thompson/familial')
```

## Usage

The `center.test()` function performs a test of centers, with the default being the Huber family of centers.
```{r, example, message = F} 
library(familial)
set.seed(1)

# One-sample test with point null
x <- MASS::galaxies
center.test(x, mu = 21000)

# One-sample test with interval null
center.test(x, mu = c(20500, 21500))

# Two-sample test
x <- MASS::cabbages[MASS::cabbages$Cult == 'c39', 'HeadWt']
y <- MASS::cabbages[MASS::cabbages$Cult == 'c52', 'HeadWt']
center.test(x, y)

# Two-sample paired directional test
x <- MASS::anorexia[MASS::anorexia$Treat == 'FT', 'Postwt']
y <- MASS::anorexia[MASS::anorexia$Treat == 'FT', 'Prewt']
center.test(x, y, paired = T, alternative = 'greater')
```

## Documentation

See the package [vignette](vignette.pdf) or [reference manual](familial_0.1.0.pdf).

<!-- See the package [vignette](https://CRAN.R-project.org/package=familial/vignettes/vignette.html) or [reference manual](https://CRAN.R-project.org/package=familial/familial.pdf). -->
